---
title: "Church Demographic Profile"
#output: hrbrthemes::ipsum_pdf
output: pdf_document
latex_engine: xelatex
params:
  mapSize: "75%"
  mapAlign: "left"
  radius: "one"
  year: 2021
  rowNum: 1
  df: vbc
  placeOfBirth: FALSE
  placeOfBirthFilter: 0
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,  dev = "cairo_pdf")
library(hrbrthemes)
library(ggplot2)
library(Cairo)
library(extrafont)
extrafont::loadfonts()
set_flextable_defaults(
    font.size = 9, theme_fun = theme_vanilla,
    padding = 0,
    background.color = "#FFFFFF",
    fonts_ignore=TRUE)

# TODO Add better comparisons. Specifically on housing expenses/values.
# TODO Develop a poverty index of sorts, to better quantify the pressure of poverty on a community. Interesting vars (B19080,B19081,B25095,C17002)
# TODO Capture variables on speaking languages and origins. (B05006)
```
This data draws on an approximately 5 mile radius from Village Baptist Church.

# Overview of Race by Asian-Alone
```{r Race by Asian Alone, message=FALSE, dev="cairo_pdf"}
displayTable(params$df,varBase = "B02015",pdf=TRUE,usCompare = FALSE,dataRowNum = 1,summaryLevels = 2)
```

# Place of Birth by Foreign-Born
The United States is a nation of immigrants. As such, it can be useful to understand where the foreign-born population in your community is located.

```{r Place of Birth Table:Foreign Born, message=FALSE, dev="cairo_pdf"}
displayTable(params$df,varBase = "B05006",pdf=TRUE,usCompare = FALSE,dataRowNum = 1,summaryLevels = 2,pctFilter = 0.005)
```

We are particularly interested in seeing where people live who were born in Vietnam. 
```{r Place of Birth Map:Foreign Born, message=FALSE, dev="cairo_pdf"}
  map_census_data(geography="tract",
                dataset = params$df$data[[params$rowNum]]$df,
                year=2021,
                coords = params$df$churchInfo[params$rowNum,],
                varBase = "B05006",
                varNum = params$placeOfBirthFilter,
                varSummaryNum = 1,
                radiusAreaOnly = FALSE, 
                filterRadius=5,
                MapTitle = "",
                LegendTitle = "",
                interactive = FALSE,
                dispPerc = FALSE)
```
